Galactic Chemical Evolution Models Favor an Extended Type Ia Supernova Delay-Time Distribution

### 2024-03-20 src/tex/ms.tex
[showyourwork] overleaf sync

The best-scoring model uses the two-infall SFH with the plateau ($W=0.3$ Gyr) DTD. Figure \ref{fig:ofe-feh-twoinfall} plots the stellar [O/Fe]--[Fe/H] abundances from this model in the inner and outer Galaxy. The model distributions in the outer Galaxy closely match the APOGEE distributions. Stellar populations at the high-[Fe/H] end of the distribution have migrated from inner regions.

@@

Table \ref{tab:results} indicates combinations of SFH and DTD which are more or less favorable in certain regimes. However, we emphasize that we do not use these scores to fit the models to the data. Much of the variation between models can be explained by the choice of SFH: in general, the two-infall models rank the highest across all diagnostics, while the late-burst models generally compare poorly to the rest. This is different for the bimodality test, where the early-burst SFH is able to produce a bimodal [O/Fe] distribution for all DTDs, as discussed in Section \ref{sec:bimodality}. The choice of DTD has a clear effect on the model scores, and this effect is similar across the diagnostics. The models which perform the best (most \yes's and fewest \no's) are those with the exponential and plateau DTDs. It is interesting that the best-performing DTD overall in our scoring system, the exponential DTD with $\tau=1.5$ Gyr, is intermediate among all our DTDs in terms of the number of prompt SNe Ia and median delay time, but has the {\it lowest} rate at very long delay times ($\sim10$ Gyr). Models with a large fraction of prompt SNe Ia, such as the power-law and two-population DTDs, fare quite poorly, with the steepest power-law ($\alpha=-1.4$) ending up in the bottom tirtile across the board for most of our SFHs. If we remove the two-infall models altogether due to their dominance in the scoring system, our results do not qualitatively change.

% The best DTDs for each SFH
The main takeaway from Table \ref{tab:results} is that while certain DTDs tend to score better than others, there is no single DTD which always produces optimal agreement with the data across all diagnostics. Depending on the choice of SFH, different DTDs will succeed at reproducing different observables. If an inside-out or late-burst SFH is chosen, the exponential DTDs with either timescale will produce optimal agreement with all observables. However, the picture gets less clear-cut when looking at the other SFHs. If the early-burst SFH is chosen, the DTDs with many prompt SNe Ia (power-law and two-population) produce the best agreement with the APOGEE MDFs but the {\it worst} agreement with the other observables. Finally, for the two-infall SFH, any DTD other than the power-law and two-population scores well, but the best-scoring DTD varies between diagnostics. In general, too many prompt SNe Ia worsens agreement with observations across all our SFHs, while too few prompt SNe Ia can also worsen agreement for some SFHs and observables.

>>>>

The best-scoring model uses the two-infall SFH with the plateau ($W=0.3$ Gyr) DTD. Figure \ref{fig:ofe-feh-twoinfall} plots the stellar [O/Fe]--[Fe/H] abundances from this model in the inner and outer Galaxy. The model distributions in the outer Galaxy closely match the APOGEE distributions. Stellar populations at the high-[Fe/H] end of the distribution have migrated from inner regions. [More discussion about Figure \ref{fig:ofe-feh-twoinfall}].

@@

% Table \ref{tab:results} indicates combinations of SFH and DTD which are more or less favorable in certain regimes. However, we emphasize that we do not use these scores to fit the models to the data. 
Much of the variation between models can be explained by the choice of SFH. In general, the two-infall models rank the highest across all diagnostics, while the late-burst models compare poorly to the rest. If we remove the two-infall models altogether due to their dominance in the scoring system, our results do not qualitatively change. The SFHs rank differently in the bimodality test, where the early-burst SFH is able to produce a bimodal [O/Fe] distribution for all DTDs, as discussed in Section \ref{sec:bimodality}.

The choice of DTD has a clear effect on the model scores, and this effect is similar across the diagnostics. The models which perform the best (most \yes's and fewest \no's) are those with the exponential and plateau DTDs. It is interesting that the best-scoring DTD overall, the exponential DTD with $\tau=1.5$ Gyr, is intermediate among all our DTDs in terms of the number of prompt SNe Ia and median delay time, but has the {\it lowest} rate at very long delay times ($\sim10$ Gyr). Models with a large fraction of prompt SNe Ia, such as the power-law and two-population DTDs, fare quite poorly, with the steepest power-law ($\alpha=-1.4$) ending up in the bottom tirtile across the board for most of our SFHs.

% The best DTDs for each SFH
% The main takeaway from Table \ref{tab:results} is that 
While certain DTDs tend to score better than others, there is no single DTD which always produces optimal agreement with the data across all diagnostics. Depending on the choice of SFH, different DTDs will succeed at reproducing different observables. If an inside-out or late-burst SFH is chosen, the exponential DTDs with either timescale will produce optimal agreement with all observables. However, the picture becomes less clear when looking at the other SFHs. If the early-burst SFH is chosen, the DTDs with many prompt SNe Ia (power-law and two-population) produce the best agreement with the APOGEE MDFs but the {\it worst} agreement with the other observables. Finally, for the two-infall SFH, any DTD other than the power-law and two-population scores well, but the best-scoring DTD varies between diagnostics. In general, too many prompt SNe Ia worsens agreement with observations across all our SFHs, while too few prompt SNe Ia can also worsen agreement for some SFHs and observables.
